---
title: <center> <h1>Crypto Bandwagon BackTester</h1> </center>
output: html_document
runtime: shiny
---
***
<center> 
**If you like it please donate to upgrade hosting**  
**Bitcoin/Bitcoin Cash:** 16f5MKogAhddR63b2zZEFx7g5yLB2oyV76   
**ERC-20:** 0x0134cfe4c693c0ae4287caa235dc540953ffbc16  
**Litecoin:** LdVZXYgF3qhGEkHKJGJ1n3gyK4w8GHvRBq  
**NANO:** xrb_39tq7hidqkknoux8sps9q7e49eky4prnaauqge8tj8yfpexmmqcx4xafn4zk  

Description of model is at the bottom of this page.  
Let me know any feature requests by filing an issue: https://github.com/MR1991/CryptoBandwagon/  
  
***


```{r setup, include = FALSE}

    knitr::opts_chunk$set(echo = FALSE)
    options(scipen = 999)
    
```

```{r libraries, include = FALSE, warning = FALSE}

    library(data.table) # fread
    library(bit64)      # integer64 type
    library(tidyr)      # data wrangling
    library(dplyr)      # data wrangling
    library(lubridate)  # dates
    library(magrittr)   # pipelines
    library(knitr)      # kable
    library(ggplot2)    # plotting
    library(plotly)     # dynamic plots

```

```{r Load data, cache = FALSE, warning = FALSE}

    # We base our analyse on a dataset obtained from Coinmarketcap.com through the 'crypto' package. 
    # https://cran.r-project.org/web/packages/crypto/index.html

    # Production dataset
    df_marketdata      <- as.data.frame(fread("Data/Crypto-Markets.csv", showProgress = FALSE)) 
    df_marketdata$date %<>% as.Date("%Y-%m-%d")
    df_marketdata$low  %<>% as.numeric()

```

```{r Check for data quality, warning = FALSE}

    # Cleaning dataset. Check whether there are any NA's or strange values in the data that we obtained from CMC. 

    marketdata_NA <- df_marketdata %>%
                     select_if(function(x) any(is.na(x))) %>% 
                     colnames()

    # one close value of coin EMV is 0, changing this to the low value of that date.
    df_marketdata[df_marketdata$close == 0, "close"] <- 0.00000000000001 

```


```{r Cleaning duplicates}

    # Check for duplicate rows in the data
    # Some coins have duplicate rows, for now we exclude these completely.
    
    duplicates <- df_marketdata %>% 
                  group_by(name, date) %>% 
                  summarise(count = n()) %>% 
                  filter(count != 1)

    df_marketdata %<>% 
        filter(!name %in% duplicates$name)

```

```{r Add historical rank}

    # Calculate % gains per day.
    df_marketdata %<>% 
        arrange(ranknow, date) %>% 
        group_by(ranknow) %>%
        mutate(prevclose = dplyr::lag(close, n = 1, default = NA)) %>%
        mutate(percchg = (close - prevclose) / prevclose * 100) %>%
        arrange(ranknow, desc(date))
    
    # Calculate rank for each day in the past. 
    df_marketdata %<>%
        group_by(date) %>% 
        mutate(rankthen = dense_rank(desc(market)))

```


```{r Model}  
        
    run_model <- function(SDate, EDate, MaxRank, HPeriod, CoinsPeriod, StopLoss, Initial, Reinvest, HoldInBTC){
            
        # Calculate % gain over previous period and actual gain over next period. Not adjusted for stoploss.
        df_modeldata <- df_marketdata %>% 
                        filter(date >= SDate - HPeriod & date <= EDate + HPeriod,
                               rankthen <= MaxRank) %>%
                        arrange(ranknow, date) %>% 
                        group_by(ranknow) %>%
                        mutate(prevclose = lag(close, n = HPeriod, default = NA))   %>%
                        mutate(prevChg   = (close - prevclose) / prevclose * 100) %>%
                        mutate(nextclose = lead(close, n = HPeriod, default = NA))  %>%
                        mutate(nextChg   = (nextclose - close) / close * 100)      %>%
                        arrange(ranknow, desc(date))
    
        # Find lowest point during next period
        df_modeldata %<>%           
            arrange(ranknow, date) %>% 
            group_by(ranknow) %>% 
            mutate(periodlow = Inf)
        
        # Dynamic loop to grab lowest point
        for (i in 1:HPeriod) {
            df_modeldata %<>% 
                mutate(leadlow = dplyr::lead(low, n = i, default = NA),
                       periodlow = ifelse(leadlow <= periodlow, leadlow, periodlow)) 
        }
        
        # Arrange dataframe again, remove unnecessary column, calculate lowest point in %, 
        # compare to stoploss and adjust.
        df_modeldata %<>%  
            arrange(ranknow, desc(date)) %>%
            select(-leadlow) %>%
            mutate(nextLow = (periodlow - close)/close * 100) %>%
            mutate(return = ifelse(nextLow <= -StopLoss * 100,  -StopLoss * 100, nextChg))   # applying stoploss
        
        # Slice off begin and end of dataframe
        df_modeldata %<>% 
            filter(date >= SDate & date <= EDate)
    
        ## Selector module}
        # Let a model select x coins per period based on the criteria and put those rows 
        # in a dataframe for result analysis
        
        df_selected = list()
    
        df_modeldata %<>% 
            arrange(desc(prevChg)) %>%
            filter(!is.na(nextclose))
        
        SwitchDates <- seq(SDate, EDate, by = paste0(HPeriod, " days"))
        
        for (i in SwitchDates){
            df_selected[[i]] <- df_modeldata %>% 
                                filter(date == i) %>% 
                                ungroup() %>% 
                                slice(1:CoinsPeriod)
        }
        
        df_selected <- bind_rows(df_selected)
        df_selected$returnfiat <- df_selected$return / 100 * Initial / CoinsPeriod
        
        # module to enable reinvest
        # if (input$Reinvest == FALSE){
        #     df_selected$returnfiat <- df_selected$return / 100 * Initial / CoinsPeriod
        # } else {
        #     
        # 
        # }
        # 
        return(df_selected)
    }

```

```{r Generate Test data, include = FALSE}

    # # INPUTS
    # # start and end date of analysis
    # SDate <- as.Date("2017-10-01")
    # EDate <- as.Date("2017-12-31")
    # 
    # # max rank of coins to be selected
    # MaxRank <- 150
    # 
    # # period between switches
    # HPeriod <- 3 # in days
    # SwitchDates <- seq(SDate, EDate, by = paste0(HPeriod, " days"))
    # 
    # # number of coins to be selected per period
    # CoinsPeriod <- 2
    # 
    # # percentage drop to sell
    # StopLoss <- 0.25 * 100  # stoploss percentage
    # 
    # # initial investment, total so its split per coin
    # Initial <- 500 # in dollars
    # 
    # # reinvest or not
    # Reinvest <- TRUE # boolean
    # 
    # # proceeds in BTC or cash
    # HoldInBTC <- FALSE # boolean
    # 
    # # (optional inputs)
    # # duration below sellpoint
    # # min % gain last week
    
     # test_data <- run_model(SDate, EDate, MaxRank, HPeriod, CoinsPeriod, StopLoss, Initial, Reinvest, HoldInBTC)


```


```{r Model Graph}

    inputPanel(
        dateRangeInput("dateRange", 
                       label = "Date range:",
                       start = min("2017-10-01", na.rm = TRUE), 
                       end = max(df_marketdata$date, na.rm = TRUE)),
        
        numericInput("MaxRank"    , "Max rank of selected coins:", value = 150),
        numericInput("HPeriod"    , "Length of period in days:" , value = 7),
        numericInput("CoinsPeriod", "# of coins per period:"    , value = 2),
        numericInput("Initial"    , "Starting funds ($):"           , value = 500),
        numericInput("StopLoss"   , "Maximum loss:"            , value = 0.25)
       # selectInput ("Reinvest"   , "Reinvest profits"          , list(`Reinvest`  = c("True","False"))), # in dev
       # selectInput ("HoldInBTC"  , "Holding earnings in BTC"   , list(`HoldInBTC` = c("True","False")))# in dev
    )

    activeDataset <- reactive({
        run_model(input$dateRange[1],
                  input$dateRange[2], 
                  input$MaxRank, 
                  input$HPeriod, 
                  input$CoinsPeriod, 
                  input$StopLoss, 
                  input$Initial, 
                  input$Reinvest == FALSE, 
                  input$HoldInBTC == FALSE) 
    })
    
  # Graph showing return per coin, including mention of period
    renderPlot({
        p <- activeDataset() %>% 
             group_by(date) %>% 
             summarize(returnfiat = sum(returnfiat)) %>%
             mutate(returnfiattime = cumsum(returnfiat),
                    pos = returnfiat >= 0)
        
        ymin <- floor(min(p$returnfiattime, p$returnfiat) / 100) * 100       
        ymax <- round(max(p$returnfiattime, p$returnfiat) / 100) * 100

        theme_set(theme_bw())
        q <- ggplot(p)+
                geom_bar(aes(x = date, y = returnfiat, fill = pos), stat = "identity")+
                geom_line(aes(x= date, y = returnfiattime), stat="identity")+
                scale_fill_manual(values = c('red', 'darkgreen')) +
               # scale_x_discrete(labels = "%b %d", breaks = paste0(input$HPeriod, " days")) +
                scale_y_continuous(breaks = seq(ymin,ymax, by = (ymax-ymin) / 10))+
                labs(title = "Return in dollars", x = "Date", y = "$")+
                guides(fill = FALSE)+
                theme(panel.background = element_rect(fill = 'white', colour = 'black'),
                      panel.grid.major = element_line(colour = 'grey'),
                      panel.grid.minor = element_line(colour = 'grey'))
        q
    })

   # div(style='height:600px; overflow-y: scroll',
    renderTable({
        t <- activeDataset() %>%
             select(symbol, name, date, prevclose, close, nextclose, prevChg, nextChg, nextLow, return) %>%
             mutate(date = format(date,'%Y/%m/%d')) %>%
             arrange(desc(date))
        names(t) <- c("Symbol", "Name", "Date", "Prev.Close ($)", "Close ($)", "Next.Close ($)", "%Prev", "%Next", "%Next.Low", "%Return")
        print(t)
    })
  #  )
  
    
```
</center>

### Input

Name | Description
------------------ | -------------------------------------------------
Date Range | The period for which you want to test your strategy, 
Max rank of selected coins | The maximum rank a coin is allowed to have for you to select it for a given period.
Length of period in days | Number of days you hold on to your coins before reevaluating.
'#' of coins per period | Number of coins you buy per period.
Starting funds ($) | Initial investment spread out over the number of coins. Current model uses this each period.
Maximum loss: | Stop loss level. Default is -25% of last close. If a coin goes under this level during a period your holdings in this coin are 'sold' until next period.

#### Output  
  
**Graph**  
The bars in the graph show the gain/loss in $ per period summed for your coins held during this period.  
The line in the graph shows your cumulative gain/loss in $ for the selected date range.

**Table**

Name | Description
------------------ | -------------------------------------------------
Symbol | Tracker symbol on CMC
Date | Date this coin was purchased
Prev.Close | Closing price at the start of the period prior to 'Date'
Close | Closing price at the mentioned date
Next.Close | Closing prive at the end of the period following 'Date'
%Prev | Percentage change over the period prior to 'Date'
%Next | Percentage change over the period following 'Date'
%Next.Low | Minimum percentage change reached during the period following 'Date'
%Return | Resulting return with this coin.



